---
title: "Registered Voters and Polling Stations in Somaliland for the 2024 Presidential and National Party Elections"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
---

```{r setup, include=FALSE}
# Load libraries
library(flexdashboard)
library(tidyverse)
library(plotly) 
library(sf)
library(gt)
library(leaflet)
library(leaflet.extras)
library(scales)
```

```{r load_data, include = FALSE}
# Load datasets
sladmin1 <- st_read("F:/OneDrive/Mukhtarabdi.com/R-DataVis/Somaliland_Elections_2024_Dashboard/data/Somaliland-Admin1.shp", quiet = TRUE)
voters <- read_csv("F:/OneDrive/Mukhtarabdi.com/R-DataVis/Somaliland_Elections_2024_Dashboard/data/total-registered-voters-per-district-2017-2024.csv")
```

## Column 1

### **Total Registered Voters for 2024 Elections**

```{r}
tot_regis_2024 <- sum(voters$`2024`, na.rm = TRUE)
valueBox(
  value = scales::comma(tot_regis_2024),    
  caption = "Total Registered Voters",        
  color = "blue"                                
)

```


### **Total Polling Stations by District**

```{r}
voters %>% 
  select(Region, District, tot_poll_stations_2024) %>% 
  gt() %>% 
  opt_interactive(use_filters = TRUE) %>% 
  tab_options(table.font.size=12) %>% 
   cols_label(Region = "Regions",
              District = "Electoral District",
              tot_poll_stations_2024 = "Total Polling Stations") %>%
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_column_labels(everything())) %>%
  fmt_number(columns = vars(tot_poll_stations_2024),
             decimals = 0) %>%
  data_color(columns = vars(tot_poll_stations_2024),
             colors = scales::col_numeric(palette = c("lightblue", "darkblue"),
                                          domain = NULL))
```

### **About the Dashboard** {data-height="150"}

This dashboard provides insights on **Somaliland's** presidential, and national party elections scheduled for November 13th, 2024. 

**Designer**: Mukhtar Abdi

- **Contact**: [mukhtarj.abdi@gmail.com](mailto:mukhtarj.abdi@gmail.com)
- **LinkedIn**: [https://www.linkedin.com/in/mukhtaabdi/]

**Data source**: Somaliland National Electoral Commision (SLNEC)

## Column 2

### **Total polling stations in 2024**

```{r}
tot_pol_2021 <- sum(voters$tot_poll_stations_2024)
valueBox(
  value = scales::comma(tot_pol_2021),
  caption = "Total Polling Stations",
  color = "blue"                                
)

```

### **Total Registered Voters by Electoral District**

```{r}
# Produce static map using ggplot2
plt1 <- voters %>% 
  drop_na() %>% 
  arrange(desc(`2024`)) %>% 
  ggplot() +
  geom_col(aes(x = `2024`, y = reorder(District, `2024`)), 
           fill = "blue") + 
  scale_x_continuous(labels = function(x) format(x, scientific = FALSE, big.mark = ",")) +
  labs(x = "Total Registered Voters", y = "Electoral District") +
  theme_minimal()

# Make the chart interactive using plotly
ggplotly(plt1)
```

## column 3

### **Electoral districts in Somaliland**

```{r}
# Get electoral districts
elecdist <- voters %>% 
  select(Region, long_X, Lat_Y, District) %>%
  drop_na() %>% 
  st_as_sf(coords = c(2:3), crs = 4326)
```


```{r}
leaflet() %>%
  addTiles() %>%
  setView(lng = 46 , lat = 9.8, zoom = 7) %>% 
  leaflet.extras::addResetMapButton() %>% 
  # Add polygon data to map
  addPolygons(data = sladmin1, 
              color = "black", 
              opacity = 0.5,
              weight = 2,
              fill = NA,
              label = ~Region_nam) %>% 
  #Add point data to the map
  addCircleMarkers(data = elecdist,
                   radius = 6,
                   color = "blue",
                   fillColor = "blue",
                   fillOpacity = 1,
                   stroke = FALSE,
                   popup = ~paste("<b>Electoral District:</b>", District)) %>% 
# Add labels to point data
  addLabelOnlyMarkers(data = elecdist,
                      label = ~District,
                      labelOptions = labelOptions(noHide = TRUE,
                                                   direction = 'top',
                                                   textsize = "8px",
                                                   opacity = 0.8)
                      )

  
```

### **Total Registered Voters by Region**

```{r}
plt2 <- voters %>%
  drop_na() %>% 
  group_by(Region) %>%
  summarise(
    total_2017 = sum(`2017`, na.rm = TRUE),
    total_2021 = sum(`2021`, na.rm = TRUE),
    total_2024 = sum(`2024`, na.rm = TRUE)
  ) %>%
  arrange(desc(total_2024)) %>%   # Arrange data from maximum to minimum for 2024
  ggplot() +
  geom_col(aes(x = reorder(Region, -total_2024), y = total_2024), # Corrected y and reordered x
           fill = "blue") +
  scale_y_continuous(labels = function(y) format(y, scientific = FALSE, big.mark = ",")) +
  labs(x = "Region", y = "Total Registered Voters") +
  theme_minimal()

# Make the chart interactive using plotly
ggplotly(plt2)


```
